---
title: "PM566 HW5"
author: "Sherry Shen"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
---
<br>

This is my PM566 HW5 website.

```{r setup, message=FALSE, echo=FALSE, warning=FALSE}
library(data.table)
library(tidyverse)
library(dplyr)
library(plotly)
library(DT)
library(knitr)

# Initialize code chunk options
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  eval=TRUE,
  echo = TRUE,
  cache = FALSE,
  fig.width = 7, 
  fig.align = 'center',
  fig.asp = 0.618,
  out.width = "700px")
```

```{r load-data, echo=FALSE}
source("process_HW5.R")
```

<br>

# Plot of number of new cases vs date outside of China

```{r}
#remove data that do not have number of case record
newcase_data=new[-(which(is.na(new$case_in_country))),c(2,3,5)]
dim(newcase_data)

#compute new cases for each country
a=as.data.frame(table(newcase_data$`reporting date`))

a %>% 
  plot_ly(x = ~Var1, y = ~Freq, type = "scatter", mode = "lines") %>%
  layout(xaxis = list(title="Date"), yaxis = list(title="Number of cases"), title = "Number of new cases outside of China")
```

From the plot we can see that, at the beginning 2 months of the COVID-29 pandemic, number of new cases outside of China is increasing as time increase. It can tell us that the virus has already been spread out around the world. 
<br>

# Evaluate the length of incubation period. 

```{r}
new %>% plot_ly(y = ~exposure_start, type = "box", name = "Exposure Start Date") %>% 
  add_trace(y = ~symptom_onset, type = "box", name = "Symptom_onset Date")

```

We can see there is around 14 days between the median of symptom onset date and exposure start date, which is the same as the incubation period of the COVID-19 virus. 

<br>

# Plot of the COVID cases outside of China

```{r}
b=as.data.frame(table(newcase_data$country))
b %>%
  group_by(Var1) %>%
  plot_ly(labels = ~ Var1, values = ~Freq, textposition = "inside") %>%
  add_pie(hole = 0.6) %>%
  layout(title = "Donus Chatrt of total numbe cases outside of China")
        
```

We can see that, countries who have large number of new cases are those located around China. This can well explain that China might be the original place of the COVID-19 virus. 

<br>
<br>
